---
title: "Animation: Building the Binomial Recursively"
author: "Joyce Robbins"
date: "3/5/2019"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

```{r make_df}
library(tidyverse)
library(gganimate)
make_df <- function(n) {
  
    all <- tibble(
    x = 0:n,
    size = paste("n =", n),
    prob = dbinom(0:n, n, .5),
    stage = paste("n =", n),
    group = 1:(n + 1),
    fill = "p(x) n trials",
    column = "n"
  )
  
  heads <- tibble(
    x = 1:(n + 1),
    size = paste("n =", n + 1),
    prob = dbinom(0:n, n, .5) / 2,
    stage = paste("n =", n, "transitioning to n =", n + 1),
    group = 1:(n + 1),
    fill = "p(x-1) n-1 trials",
    column = "n - 1"
  )
  
  tails <- tibble(
    x = 0:n,
    size = paste("n =", n + 1),
    prob = dbinom(0:n, n, .5) / 2,
    stage = paste("n =", n, "transitioning to n =", n + 1),
    group = 1:(n + 1),
    fill = "p(x) n-1 trials",
    column = "n - 1"
  )
  
  rbind(all, heads, tails) %>%
    mutate(
      fill = fct_relevel(
        factor(fill),
        "p(x-1) n-1 trials",
        "p(x) n-1 trials",
        "p(x) n trials"
      ),
      column = fct_relevel(column, "n - 1")
    )
}

num <- 12

df <- purrr::map_dfr(1:num, make_df) %>% 
  mutate(size = factor(size, levels = unique(size)),
         stage = factor(stage, levels = unique(stage)))
```


If $p=.5$, half of the sum of $P(X=x) + (P = x-1)$ for $n$ trials gives us $(P = x)$ for $n+1$ trials. (For a more thorough explanation on why this is true, see [Understandng the Binomial Recursively](https://jtr13.github.io/website/Binomial.html).)

That is,

$$\frac{b(x; n, .5)}{2} + \frac{b(x-1; n, .5)}{2} = b(x; n+1, .5)$$

This animation demonstrates this recursion visually for $n = 2$ to $n = `r num`$: the red and blue represent half of the probabilities for $n$ trials, while the purple bars represent the binomial probabilities for $n+1$ trials. (Note that the blue bars are half the height of the purple bars--the $x's$ are the same so the purple bars shrink into the blue ones for $n+1$ trials.)


```{r animation}
g <-
  ggplot(df, aes(x, prob, fill = fill, group = group)) +
  geom_col() + 
  scale_fill_manual(values = c("red", "blue", "purple")) +
  labs(title = 'Binomial Distribution: {closest_state}', y = 'probability') + theme(legend.position = "bottom", 
                                                                                    legend.title = element_blank())

anim <- 
  g + transition_states(stage) + exit_recolor(fill = "purple")

animate(anim, fps = 5)

```

Animations can be hard to digest, so here's a static version:


```{r static, fig.height = 10}
df %>% 
  mutate(sizenum = parse_number(as.character(size))) %>% 
  filter(sizenum != 1, sizenum != max(sizenum)) %>% 
  ggplot(aes(x, prob, fill = fill)) +
  geom_col() +
  scale_x_continuous(breaks = 0:max(df$x)) +
  scale_fill_manual(values = c("red", "blue", "purple")) +
  facet_grid(size ~ column) + 
  ggtitle("Binomial distribution") +
  theme(legend.position = "bottom", 
        legend.title = element_blank())
```

[Back](https://jtr13.github.io/website/)
